\documentclass[pdflatex,sn-mathphys,Numbered]{sn-jnl}

\usepackage{graphicx}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{amsthm}
\usepackage{mathrsfs}
\usepackage[title]{appendix}
\usepackage{xcolor}
\usepackage{textcomp}
\usepackage{manyfoot}
\usepackage{booktabs}

\theoremstyle{thmstyleone}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{question}[theorem]{Question}
\theoremstyle{thmstyletwo}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{convention}[theorem]{Convention}

\newcommand{\MH}{\mathrm{MH}}
\newcommand{\PH}{\mathrm{PH}}
\newcommand{\RH}{\mathrm{RH}}
\newcommand{\MC}{\mathrm{MC}}
\newcommand{\QQ}{\mathbb{Q}}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\dwl}{\mathrm{DWL}}
\newcommand{\dwlp}{\mathrm{DWL}^{+}}

\input{tables_generated}

\raggedbottom

\begin{document}

\title[Directed Weisfeiler--Leman refinement and digraph homology]{Directed
Weisfeiler--Leman refinement versus the invariants of the magnitude--path
spectral sequence}

\author*[1]{\fnm{Krishna} \sur{Harish}}\email{krishnaharish2009@gmail.com}

\affil*[1]{\orgname{Elkins High School}, \orgaddress{\city{Missouri City},
\state{TX}, \country{USA}}}

\abstract{Magnitude homology, path homology, and reachability homology are
organized by the magnitude--path spectral sequence (MPSS) of a digraph:
magnitude homology is its first page, path homology sits on an axis of its
second page, and reachability homology is its target. Directed
Weisfeiler--Leman color refinement (D-WL) is the standard upper bound for
the separating power of message-passing neural networks on digraphs. We
give the first systematic comparison of the two families, the directed
counterpart of the known placement of persistent homology in the undirected
WL hierarchy. First, the families are incomparable: the directed $4$-cycle
and the disjoint union of two digons are D-WL-equivalent but are separated
by all three homology theories, while an in-star and a directed path of
length two are separated by D-WL although all three homology theories
agree; by exhaustive exact computation over all $1{,}550{,}786$ digraphs
on three to six vertices, both witnesses are minimal. The incomparability
persists for the reciprocity-aware refinement $\dwlp$, with minimal witness
exactly at six vertices, and indeed magnitude and path homology separate
six-vertex pairs on which $\dwlp$ \emph{and} the distance multiset agree,
by detecting asymmetric round-trip structure that survives geodesic
filling. Second, the bigraded rank function of the first
page of the MPSS does not determine the second page: any two digraphs in
which every vertex is a source or a sink and which share vertex and arc
counts have isomorphic magnitude homology in every bidegree, yet path
homology recovers the Betti numbers of the underlying graph; on at most
four vertices, exhaustively, no such rank-equal pair with distinct path
homology exists, so five vertices is minimal. Third, magnitude homology
separates D-WL-equivalent digraphs with equal directed-distance multisets,
so its excess power over refinement is not distance bookkeeping; the
minimal witnesses (five vertices, none at four or fewer) pit the symmetric
digraph of the $5$-cycle against non-circulant partial orientations. A
component-counting description of degree-two magnitude homology identifies
the mechanism behind all of these separations: $\MH_{2,2}$ decomposes into
a refinement-visible term minus a distance term, while diagonal summands
of $\MH_{2,l}$ count directed round trips (girth and reciprocity
structure), exactly the correlation that refinement and distances discard.
All homology ranks are computed exactly over $\QQ$; we release the
accompanying toolkit, to our knowledge the first for magnitude, path, and
reachability homology of general digraphs in a single exact-arithmetic
framework.}

\keywords{magnitude homology, path homology, reachability homology,
Weisfeiler--Leman, digraphs, graph neural networks, expressivity}

\maketitle

\section{Introduction}\label{sec:intro}

Two research communities have, largely independently, built graph-level
invariants of directed graphs.

On the applied-topology side, a family of homology theories for digraphs
has crystallized over the past decade: the path homology of Grigor'yan,
Lin, Muranov, and Yau \cite{GLMY12,GLMY14}, sensitive to the composability
of arcs; the magnitude homology of Hepworth--Willerton and
Leinster--Shulman \cite{HepworthWillerton17,LeinsterShulman21},
categorifying Leinster's magnitude \cite{Leinster13} and defined for any
digraph through its shortest-path quasimetric; and the reachability
homology of Hepworth and Roff \cite{HepworthRoffIMRN}. These are not
independent inventions: Asao \cite{Asao23} constructed a spectral sequence,
the magnitude--path spectral sequence (MPSS), whose first page is magnitude
homology and whose second page contains path homology, and Hepworth and
Roff \cite{HepworthRoff24,HepworthRoffIMRN} showed that it converges to
reachability homology. The MPSS thus organizes the three theories into
pages of a single object.

On the machine-learning side, the separating power of message-passing
graph neural networks is bounded by Weisfeiler--Leman color refinement
\cite{Xu19,Morris19}; for digraphs the relevant procedure is directed
color refinement (here D-WL), which refines vertex colors by the multisets
of in-neighbor and out-neighbor colors separately \cite{GroheColRef,
Jaume19}. For undirected graphs, the relationship between persistent
homology and the WL hierarchy has been worked out in detail by Ballester
and Rieck \cite{BallesterRieck25}, but that study is explicitly restricted
to undirected graphs, and we are not aware of any placement of the
digraph homology theories above relative to directed refinement. The
closest prior results are single-invariant comparisons within topology,
such as the digraph pair of Chaplin, Harrington, and Tillmann
\cite{ChaplinHT24} distinguishing path homology from directed flag
complex homology.

This paper carries out the comparison. Our contributions are as follows;
throughout, $\MH$, $\PH$, $\RH$ denote magnitude, path, and reachability
homology over $\QQ$ (Section~\ref{sec:background}), and all computational
claims are exact rational computations, exhaustive over all digraphs on
three to six vertices up to isomorphism
($16 + 218 + 9{,}608 + 1{,}540{,}944 = 1{,}550{,}786$ digraphs; the four
digraphs on at most two vertices are pairwise D-WL-distinguished, so all
minimality thresholds are meaningful).

\begin{enumerate}
\item \textbf{Incomparability} (Theorem~\ref{thm:incomparable}). D-WL and
the MPSS invariants each separate digraph pairs the other cannot. The
directed $4$-cycle versus two digons is D-WL-equivalent but separated by
each of $\MH$, $\PH$, $\RH$; an in-star versus a directed $2$-path is
D-WL-separated while all three homology theories agree. Both witnesses are
minimal by exhaustive search. For the reciprocity-aware refinement
$\dwlp$, which additionally distinguishes reciprocal neighbors, the
homology-beats-refinement witness moves to the directed $6$-cycle versus
two directed triangles, minimal at exactly six vertices
(Theorem~\ref{thm:dwlplus}).
\item \textbf{The first page does not determine the second}
(Theorem~\ref{thm:e1e2}). On the family of digraphs in which every vertex
is a source or a sink, all magnitude homology is concentrated in bidegrees
$(0,0)$ and $(1,1)$ with ranks $(n, m)$, while path homology and
reachability homology compute the Betti numbers of the underlying
undirected graph. Consequently the bigraded ranks of the $E^1$ page of the
MPSS do not determine its $E^2$ page, its limit, or even weak
connectivity. Exhaustively, no pair of digraphs on at most four vertices
with equal $\MH$ ranks has distinct path homology, so the minimal such
pair, the in-star $K_{4\to 1}$ versus $K_{2\to 2}\sqcup\ast$, lives on
five vertices.
\item \textbf{Beyond distances} (Theorem~\ref{thm:beyonddist}). Magnitude
homology ranks do not determine the directed-distance multiset
(Proposition~\ref{prop:mhnotdist}), and conversely magnitude homology
separates D-WL-equivalent pairs with \emph{equal} distance multisets:
minimally, on five vertices, the symmetric digraph of the $5$-cycle
against a partial orientation with three digons, where path homology even
distinguishes a $1$-dimensional from a $2$-dimensional hole. All such
witnesses on at most five vertices carry unequal digon counts, so the
reciprocity-aware $\dwlp$ together with distances subsumes the homology
package in that range; but at six vertices, minimally, magnitude
homology and path homology separate even
$(\dwlp,\text{distance})$-equivalent pairs
(Theorem~\ref{thm:beyonddwlp}): what they detect is asymmetric
round-trip structure modulo geodesic fillability
(Remark~\ref{rem:roundtripdepth}), which arc-typing by reciprocation
cannot see.
\item \textbf{The mechanism} (Section~\ref{sec:mech}). Building on the
degree-two basis of Ivanov and Mukoseev \cite{IvanovMukoseev24}, we give
a midpoint-graph component formula for $\MH_{2,l}$ with a short direct
proof, verified mechanically in every grading on all $9{,}842$ digraphs.
Its corollaries locate the power of magnitude homology precisely:
$\MH_{2,2} = \sum_v \deg^-(v)\deg^+(v) - \#\{d = 2\}$ is refinement data
minus distance data, while diagonal summands of $\MH_{2,l}$ count
directed round trips, a directed-girth and reciprocity signal
(cf.~\cite{AHK24}) that neither D-WL nor distances retain.
\item \textbf{The landscape} (Section~\ref{sec:landscape}). We tabulate
the partition of all $1{,}550{,}786$ digraphs on at most six vertices
induced by each invariant and each pairwise refinement relation,
including that $\PH$ does not determine $\RH$, while $\MH$ and $\PH$
jointly determine $\RH$ throughout this range
(Question~\ref{q:mhphrh}).
\item \textbf{Software.} The exact-arithmetic toolkit and all data are
released; to our knowledge it is the first public implementation
computing magnitude homology of general digraphs, and the first computing
the three MPSS-related theories in one framework. (For undirected graphs,
computations of graph magnitude homology appear in the literature, for
example via algebraic Morse theory \cite{Gu18}.)
\end{enumerate}

Beyond the specific statements, the results calibrate expectations for
homology-based features in directed graph learning: the MPSS invariants
are neither refinements nor coarsenings of the D-WL bound that governs
standard directed message passing, so they are candidates for genuinely
complementary features, but they also miss elementary refinement
information (weak connectivity is invisible to magnitude homology, while
degree profiles separate what all three theories jointly cannot). We
return to this in Section~\ref{sec:discussion}.

\subsection*{Relation to prior work}

Ballester and Rieck \cite{BallesterRieck25} place persistent homology of
filtered clique complexes within the (undirected) WL hierarchy; our
results are the directed analogue at the level of unfiltered digraph
homology theories, and Section~\ref{sec:discussion} discusses the
persistent directed extension as future work. Path-based neural
architectures \cite{Michel23} exceed 1-WL by aggregating over paths but
involve no homology. Within topology, comparisons among directed
invariants exist \cite{ChaplinHT24,CaputiMR25}, but none against
refinement hierarchies. On the computational side, Menara and Manzoni
\cite{MenaraManzoni24} analyze the parameterized complexity of Eulerian
magnitude homology of graphs; our sweep is feasible precisely because the
digraphs are small and the arithmetic is exact. The MPSS papers
\cite{Asao23,HepworthRoff24,HepworthRoffIMRN} study the algebraic
structure of the tower; our Theorem~\ref{thm:e1e2} is, to our knowledge,
the first explicit witness that the rank function of $E^1$ does not
determine $E^2$, complementing their structural results. Ivanov and
Mukoseev \cite{IvanovMukoseev24} give a complete basis for degree-two
magnitude homology of digraphs in the course of studying diagonality;
Section~\ref{sec:mech} uses an equivalent component form of that
description and draws the separating-power consequences, which are not
considered there. Persistent
magnitude homology has just been given foundations \cite{BiFengLiWu26};
its separating power is untouched and inherits our lower bounds.

\section{Background and conventions}\label{sec:background}

\begin{convention}
A \emph{digraph} $G = (V, A)$ is a finite set $V$ with a set
$A \subseteq V \times V$ of ordered pairs (\emph{arcs}) with no loops;
$n = |V|$, $m = |A|$. A \emph{digon} is a pair of reciprocal arcs
$\{(u,v),(v,u)\}$. The \emph{quasimetric} $d(x,y) \in \{0,1,2,\dots\}
\cup \{\infty\}$ is the directed shortest-path distance. All homology is
taken with coefficients in $\QQ$ and we record ranks. Digraph
enumeration is up to isomorphism via nauty \cite{nauty}; the counts
$16$, $218$, $9{,}608$ for $n = 3,4,5$ match OEIS A000273.
\end{convention}

\subsection{Directed color refinement}

D-WL iteratively refines a vertex coloring: all vertices start with a
common color, and the color of $v$ at round $t{+}1$ is the (injectively
encoded) triple of its color at round $t$, the multiset of round-$t$
colors of its in-neighbors, and the multiset of round-$t$ colors of its
out-neighbors \cite{GroheColRef,Jaume19}. Since a partition of an $n$-set
can strictly refine at most $n-1$ times, the partition stabilizes within
$n$ rounds. Two digraphs are \emph{D-WL-equivalent} if the color
histograms agree at every round (colors encoded canonically across
graphs). This is the separating power that bounds standard directed
message-passing networks, in the same way that 1-WL bounds GNNs on
undirected graphs \cite{Xu19,Morris19}.

We will also use a natural strengthening. The \emph{reciprocity-aware}
refinement $\dwlp$ refines by the triple of multisets over in-only
neighbors, out-only neighbors, and reciprocal neighbors (those joined to
$v$ by a digon). $\dwlp$ refines D-WL; it corresponds to message passing
that treats reciprocated and unreciprocated arcs as different edge types,
as many practical digraph architectures do.

\subsection{Magnitude homology}

Following \cite{HepworthWillerton17,LeinsterShulman21}, applied to the
quasimetric space $(V, d)$: for $k \ge 0$ and $l \ge 0$, the chain group
$\MC_{k,l}(G)$ is the $\QQ$-span of tuples $(x_0, \dots, x_k)$ of
vertices with $x_{i-1} \ne x_i$, $d(x_{i-1}, x_i) < \infty$, and
$\sum_{i=1}^{k} d(x_{i-1}, x_i) = l$. The differential is
$\partial = \sum_{i=1}^{k-1} (-1)^i \partial_i$, where $\partial_i$
deletes $x_i$ if $d(x_{i-1},x_i) + d(x_i,x_{i+1}) = d(x_{i-1},x_{i+1})$
and is zero on the tuple otherwise. Magnitude homology $\MH_{k,l}(G)$ is
the homology of the $l$-graded piece. Its graded Euler characteristic
recovers the magnitude power series; we use this identity as an
end-to-end correctness test of our implementation
(Section~\ref{sec:landscape}). By $\MH$ \emph{ranks} we always mean the
function $(k,l) \mapsto \dim_\QQ \MH_{k,l}$.

\subsection{Path homology}

Following \cite{GLMY12,GLMY14}: an \emph{allowed} $p$-path is a vertex
sequence $(x_0,\dots,x_p)$ each consecutive pair of which is an arc. On
the span of \emph{regular} paths (consecutive entries distinct) one takes
the alternating face boundary with irregular faces set to zero, and
$\Omega_p = \{\omega \in \operatorname{span}(A_p) : \partial\omega \in
\operatorname{span}(A_{p-1})\}$ is the space of $\partial$-invariant
allowed chains. Path homology $\PH_p(G)$ is the homology of
$(\Omega_\ast, \partial)$. We compute $\PH_0, \PH_1, \PH_2$.

\subsection{Reachability homology and the MPSS}

Reachability homology $\RH_\ast(G)$ \cite{HepworthRoffIMRN} is the
homology of the nerve of the reachability preorder ($x \preceq y$ iff
$d(x,y) < \infty$). The preorder is equivalent, as a category, to its
condensation poset (strong components ordered by reachability), so
$\RH_\ast(G)$ is the simplicial homology of the order complex of the
condensation; this is how we compute it, in dimensions $0$ to $3$.
Asao \cite{Asao23} constructs the MPSS with $E^1 \cong \MH$; the diagonal
of its $E^2$ page is reduced path homology, and the sequence converges to
reachability homology \cite{HepworthRoff24,HepworthRoffIMRN}. We only use
this organizational fact: $\MH$, $\PH$, $\RH$ are, respectively, the
first page, part of the second page, and the target of one spectral
sequence.

\subsection{Windows, and what is absolute}\label{subsec:windows}

Sweep-level computations use the windows $k \le 3$ (all $l$) for $\MH$,
$p \le 2$ for $\PH$, and dimensions $\le 3$ for $\RH$. An
\emph{inequality} of invariants observed in a window is absolute: the
full invariants differ. An \emph{equality} claim is, in general, relative
to the window and is stated as such; the star witnesses of
Theorem~\ref{thm:e1e2} are the exception, since for them the full
magnitude homology is determined by hand in all bidegrees.

\section{Structural lemmas}\label{sec:lemmas}

\begin{lemma}\label{lem:regular}
Let $G$ and $H$ be digraphs on $n$ vertices in which every vertex has
out-degree $a$ and in-degree $b$. Then $G$ and $H$ are D-WL-equivalent.
If moreover neither contains a digon, then $G$ and $H$ are
$\dwlp$-equivalent.
\end{lemma}

\begin{proof}
By induction on the round: all vertices of both digraphs share the
initial color, and if all share a color $c_t$ at round $t$, then every
vertex of either digraph has the same signature
$(c_t, \{c_t^{(b)}\}, \{c_t^{(a)}\})$, hence the same round-$(t{+}1)$
color. The histograms therefore agree at every round. In the digon-free
case the reciprocal multiset is empty for every vertex and the same
induction applies to $\dwlp$.
\end{proof}

\begin{lemma}\label{lem:mh1}
For every digraph $G$: $\MH_{0,0}(G) \cong \QQ^{n}$,
$\MH_{1,1}(G) \cong \QQ^{m}$, and $\MH_{1,l}(G) = 0$ for all $l \ge 2$.
\end{lemma}

\begin{proof}
Degree $0$ is immediate. $\MC_{1,l}$ is spanned by ordered pairs at
distance exactly $l$, and the outgoing differential vanishes on
$\MC_{1,\ast}$ (there are no interior vertices to delete). At $l = 1$
the pairs at distance $1$ are exactly the arcs and $\MC_{2,1} = 0$, so
$\MH_{1,1} \cong \QQ^m$. For $l \ge 2$, let $d(x,y) = l$ and let $z$ be
the first interior vertex of a geodesic from $x$ to $y$; then
$d(x,z) = 1$ and $d(z,y) = l - 1$, so $(x,z,y) \in \MC_{2,l}$ and
$\partial(x,z,y) = -(x,y)$. Hence the incoming differential is
surjective and $\MH_{1,l} = 0$.
\end{proof}

Thus all information carried by magnitude homology beyond the vertex and
arc counts lives in homological degrees $k \ge 2$. The next observation
identifies a family where those degrees vanish identically.

\begin{lemma}\label{lem:sourcesink}
Call a digraph \emph{layered} if every vertex is a source or a sink
(no vertex has both an incoming and an outgoing arc). If $G$ is layered
with $n$ vertices, $m$ arcs, and $c$ weakly connected components, then:
\begin{enumerate}
\item $\MC_{k,l}(G) = 0$ for all $k \ge 2$, hence $\MH(G)$ has rank $n$
at $(0,0)$, rank $m$ at $(1,1)$, and rank $0$ elsewhere;
\item $\PH_0(G) \cong \QQ^{c}$, $\PH_1(G) \cong \QQ^{\,m - n + c}$, and
$\PH_p(G) = 0$ for $p \ge 2$;
\item $\RH_0(G) \cong \QQ^{c}$, $\RH_1(G) \cong \QQ^{\,m - n + c}$, and
$\RH_p(G) = 0$ for $p \ge 2$.
\end{enumerate}
\end{lemma}

\begin{proof}
(1) A tuple $(x_0, x_1, x_2, \dots)$ in $\MC_{k,l}$ with $k \ge 2$
requires $d(x_0, x_1) < \infty$ and $d(x_1, x_2) < \infty$ with
$x_0 \ne x_1 \ne x_2$, so $x_1$ has both an incoming and an outgoing arc
(any finite positive distance requires at least one arc out of the source
and into the target), contradicting layeredness.
(2) In a layered digraph there are no allowed $2$-paths, so
$\Omega_p = 0$ for $p \ge 2$, $\Omega_1$ is spanned by the arcs, and
$\Omega_0$ by the vertices. The complex
$0 \to \QQ^m \xrightarrow{\;\partial\;} \QQ^n \to 0$ is the simplicial
chain complex of the underlying undirected graph, whose homology is
$(\QQ^c, \QQ^{m-n+c})$.
(3) Layeredness forbids digons and directed $2$-paths, so the
reachability preorder is the height-one poset whose comparabilities are
the arcs, all strong components are singletons, and the order complex is
the underlying undirected graph (its edges are the arcs; there are no
chains of length two). The claim follows as in (2).
\end{proof}

\begin{proposition}\label{prop:mhnotdist}
Magnitude homology ranks do not determine the multiset of directed
distances. Minimal witnesses exist on three vertices: the in-star
$S$ ($u \to w \leftarrow v$) and the directed path
$P$ ($u \to w \to v$) satisfy $\MH_{k,l}(S) \cong \MH_{k,l}(P)$ for all
$(k,l)$, but the finite distance multisets are $\{1,1\}$ and
$\{1,1,2\}$.
\end{proposition}

\begin{proof}
$S$ is layered, so by Lemma~\ref{lem:sourcesink} its $\MH$ is
$(3, 2)$ at bidegrees $(0,0), (1,1)$ and zero elsewhere. In $P$, the only
tuples beyond degree one are built from the composable pair
$u \to w \to v$: $\MC_{2,2}(P)$ is spanned by $(u,w,v)$, and
$\MC_{k,l}(P) = 0$ for $k \ge 3$ (nothing leaves $v$). Since
$d(u,w) + d(w,v) = 2 = d(u,v)$, we get
$\partial(u,w,v) = -(u,v) \ne 0$, so the single degree-$2$ chain is not a
cycle and it kills the single class of $\MC_{1,2}$: explicitly
$\MH_{1,2}(P) = 0$, $\MH_{2,2}(P) = 0$. Hence $\MH(P)$ also equals
$(3,2)$ at $(0,0), (1,1)$ and zero elsewhere, while $d(u,v) = 2$ appears
only in $P$.
\end{proof}

\section{Incomparability}\label{sec:incomparable}

\begin{theorem}\label{thm:incomparable}
D-WL and the family $\{\MH, \PH, \RH\}$ are incomparable in separating
power:
\begin{enumerate}
\item[(a)] The directed $4$-cycle $C_4$ and the disjoint union of two
digons $D_2 \sqcup D_2$ are D-WL-equivalent, but each of $\MH$, $\PH$,
$\RH$ separates them. No such pair exists on at most three vertices.
\item[(b)] The in-star $S$ and directed path $P$ of
Proposition~\ref{prop:mhnotdist} satisfy
$\MH(S) \cong \MH(P)$ (all bidegrees), $\PH_\ast(S) \cong \PH_\ast(P)$,
$\RH_\ast(S) \cong \RH_\ast(P)$, but D-WL separates them at round one.
\end{enumerate}
\end{theorem}

\begin{proof}
(a) Both digraphs are $1$-in-$1$-out regular on four vertices, so
Lemma~\ref{lem:regular} gives D-WL-equivalence. $\RH$: $C_4$ is strongly
connected, so its reachability preorder is equivalent to a point and
$\RH_0 = \QQ$; the union of digons has two strong components, mutually
unreachable, giving $\RH_0 = \QQ^2$. $\PH$: $\PH_0$ counts weak
components ($1$ versus $2$); moreover $\PH_1(C_4) = \QQ$ (the four arcs
support the cycle class $\sum e_i$, and $\Omega_2 = 0$ since every
allowed $2$-path $(x, x{+}1, x{+}2)$ has the non-allowed face
$(x, x{+}2)$, and these four faces are distinct, so no linear combination
cancels them), while $\PH_1(D_2 \sqcup D_2) = 0$: for a single digon on
$\{x,y\}$, $\Omega_1$ is spanned by $e_{xy}, e_{yx}$ with $\partial$ of
rank $1$, and the surviving cycle $e_{xy} + e_{yx}$ is a boundary, since
the allowed $2$-path $(x,y,x)$ is $\partial$-invariant with
$\partial(x,y,x) = e_{yx} + e_{xy}$ (the middle face $(x,x)$ is
irregular, hence zero), so $\PH_1 = 0$ for each digon and their disjoint
union.
$\MH$: the machine computation (exact over $\QQ$, Table~\ref{tab:star})
gives $\MH_{2,4}(C_4) = \QQ^4$ while $\MH_{2,4}(D_2 \sqcup D_2) = 0$;
alternatively, the digon union is the symmetric digraph of a forest-like
graph ($2 K_2$), whose magnitude homology is diagonal
\cite{HepworthWillerton17}, while $C_4$ has nonzero off-diagonal ranks.
Since a rank inequality observed in any bidegree is absolute
(Section~\ref{subsec:windows}), all three separations hold. Minimality is
the exhaustive $n = 3$ computation (Table~\ref{tab:splits}: zero D-WL
classes split by any homology at $n = 3$).

(b) Magnitude homology equality is Proposition~\ref{prop:mhnotdist}.
$\PH$: both digraphs are weakly connected with two arcs whose boundaries
are linearly independent, and in both $\Omega_2 = 0$ ($S$ has no allowed
$2$-path; in $P$ the single allowed $2$-path $(u,w,v)$ has non-allowed
face $(u,v)$), so $\PH_\ast = (\QQ, 0, 0)$ for both. $\RH$: the
reachability posets are $u, v \prec w$ (order complex: a path, hence
contractible) and $u \prec w \prec v$ with $u \prec v$ (order complex: a
$2$-simplex, contractible), so $\RH_\ast = (\QQ, 0, 0, 0)$ for both.
D-WL separates at round one: the in-degree multisets are
$\{0,0,2\}$ and $\{0,1,1\}$.
\end{proof}

\begin{table}[t]
\centering
\caption{Magnitude homology ranks of the star witnesses of
Theorem~\ref{thm:incomparable}(a) and Theorem~\ref{thm:beyonddist}
(machine-generated; nonzero bidegrees $(k,l)$ with $k \le 3$).}
\label{tab:star}
\begin{tabular}{ll}
\toprule
digraph & nonzero $\MH_{k,l}$, $k \le 3$ \\
\midrule
directed $C_4$ & \mhcfour \\
$D_2 \sqcup D_2$ & \mhdigons \\
symmetric $C_5$ (\texttt{\&DKMME?}) & \mhcirc \\
twin (\texttt{\&DKLMI?}) & \mhcirctwin \\
\bottomrule
\end{tabular}
\end{table}

The refinement $\dwlp$ sees digons, and the witness of part (a) has
digons on one side only. The incomparability nevertheless persists, one
size up, and exactly there:

\begin{theorem}\label{thm:dwlplus}
The directed $6$-cycle $C_6$ and the disjoint union $C_3 \sqcup C_3$ of
two directed triangles are $\dwlp$-equivalent, and each of $\MH$, $\PH$,
$\RH$ separates them. Exhaustively, no $\dwlp$-equivalent pair of
digraphs on at most five vertices is separated by $\MH$ (ranks
$k \le 3$) or by $\PH_{\le 2}$; hence six vertices is minimal. There,
exactly $140$ $\dwlp$-classes are split by $\MH$ and $38$ by $\PH$.
\end{theorem}

\begin{proof}
Both digraphs are digon-free and $1$-in-$1$-out regular on six vertices,
so Lemma~\ref{lem:regular} gives $\dwlp$-equivalence. Separations:
$\PH_0$ and $\RH_0$ count weak components and strong condensation
components ($1$ versus $2$ each); $\PH_\ast(C_6) = (\QQ, \QQ, 0)$ and
$\PH_\ast(C_3 \sqcup C_3) = (\QQ^2, \QQ^2, 0)$ (directed cycles have
$\PH_1 = \QQ$ \cite{GLMY12}; verified exactly by our implementation);
$\MH_{1,\ast}$ agrees (Lemma~\ref{lem:mh1}) but for instance
$\MC_{1,3}$-level distances differ and the machine gives, absolutely,
$\MH_{2,3}(C_3 \sqcup C_3) = \QQ^{6} \ne \QQ^{0} = \MH_{2,3}(C_6)$.
The exhaustive statement is a direct computation over all $9{,}842$
digraphs on $3 \le n \le 5$, together with the trivial check that the
four digraphs on $n \le 2$ are pairwise D-WL-distinguished; the $n = 6$
split counts come from the exhaustive $n = 6$ sweep
(Section~\ref{sec:landscape}).
\end{proof}

\section{The first page does not determine the second}\label{sec:e1e2}

\begin{theorem}\label{thm:e1e2}
Let $X = K_{4 \to 1}$ (four sources, one sink, all four arcs) and
$Y = K_{2\to 2} \sqcup \ast$ (complete bipartite orientation plus an
isolated vertex), both on five vertices with four arcs. Then:
\begin{enumerate}
\item[(a)] $\MH_{k,l}(X) \cong \MH_{k,l}(Y)$ for \emph{all} $(k,l)$;
\item[(b)] $\PH_\ast(X) = (\QQ, 0, 0) \ne (\QQ^2, \QQ, 0) = \PH_\ast(Y)$
and $\RH_\ast(X) = (\QQ, 0, 0, 0) \ne (\QQ^2, \QQ, 0, 0) =
\RH_\ast(Y)$.
\end{enumerate}
Hence the bigraded rank function of the first page of the MPSS determines
neither its second page nor its target, and magnitude homology does not
determine weak connectivity. Exhaustively, no pair of digraphs on at most
four vertices with equal $\MH$ ranks ($k \le 3$, all $l$) has distinct
$\PH_{\le 2}$, so five vertices is minimal.
\end{theorem}

\begin{proof}
Both digraphs are layered with $(n, m) = (5, 4)$;
Lemma~\ref{lem:sourcesink}(1) gives (a), and
Lemma~\ref{lem:sourcesink}(2),(3) give (b) with $(c, m{-}n{+}c)$ equal to
$(1, 0)$ for $X$ and $(2, 1)$ for $Y$. For the MPSS statement: the $E^1$
pages have equal bigraded ranks by (a), while the $E^2$ page contains
(reduced) path homology \cite{Asao23,HepworthRoff24} and the target is
reachability homology \cite{HepworthRoffIMRN}, both of which differ by
(b). Minimality is the exhaustive $n \le 4$ computation
(Table~\ref{tab:splits}); note that an equality of the coarser $k \le 3$
rank function is implied by equality of the full rank function, so the
exhaustive statement covers the full-rank hypothesis a fortiori.
\end{proof}

\begin{remark}
Lemma~\ref{lem:sourcesink} yields infinitely many such pairs: any two
layered digraphs with equal $(n, m)$ and different $c$ (equivalently,
different cycle rank $m - n + c$) work. On five vertices with four arcs
the pair $(X, Y)$ above is the one produced by the exhaustive search. The
phenomenon is a rank statement about $E^1$: as pages-with-differentials
the two $E^1$ terms are of course not isomorphic, and the theorem shows
the differential carries essential information, in the strongest sense
that it is not recoverable from ranks.
\end{remark}

\begin{remark}\label{rem:truncation}
The sweep also finds pairs equal at $k \le 3$ that split at $k = 4$
(for example \texttt{\&DKCJN?} versus \texttt{\&DKWMM?}, whose $\MH_{4,4}$
ranks are $4$ and $6$): window equality is genuinely weaker than full
equality, which is why Theorem~\ref{thm:e1e2} is stated with a
hand-proved full-rank witness.
\end{remark}

\section{Beyond distance multisets}\label{sec:beyonddist}

D-WL does not determine directed distances, and distances do not
determine D-WL (Table~\ref{tab:splits}, rows 3 and 4); the pair
$(C_4, D_2 \sqcup D_2)$ realizes the first failure. The sharper question
is whether the homology theories exceed D-WL \emph{only} through
distances. On at most four vertices the answer is yes: every D-WL class
split by $\MH$ or $\PH$ is already split by the distance multiset. On
five vertices the answer is no:

\begin{theorem}\label{thm:beyonddist}
There exist digraphs on five vertices, minimally, that are
D-WL-equivalent and have equal multisets of directed distances but are
separated by $\MH$ and by $\PH$. Witness (machine-verified, exact): the
symmetric digraph $\overline{C_5}$ of the $5$-cycle
(arcs $x \to x \pm 2 \bmod 5$; five digons) and its partial-orientation
twin \texttt{\&DKLMI?} (arcs
$\{02, 03, 13, 14, 21, 24, 30, 31, 40, 42\}$; three digons). Both are
$2$-in-$2$-out regular, strongly connected with all $20$ ordered
distances in $\{1, 2\}$ ($10$ each), and D-WL-equivalent by
Lemma~\ref{lem:regular}; but $\MH_{2,3}$ has ranks $10$ versus $6$,
$\MH_{3,4}$ has ranks $30$ versus $26$, and
$\PH_\ast = (\QQ, \QQ, 0)$ versus $(\QQ, 0, \QQ)$: the twin trades the
$1$-dimensional hole of $\overline{C_5}$ for a $2$-dimensional one.
Exactly three such D-WL-and-distance-equivalent split classes exist on
five vertices, listed in Appendix~\ref{app:witnesses}.
\end{theorem}

\begin{proof}
Regularity, distances, digon counts, and D-WL-equivalence are finite
checks stated in the theorem (the distance claim follows since each
vertex has two out-neighbors and reaches the remaining two vertices in
two steps; Lemma~\ref{lem:regular} applies). The homology inequalities
are exact machine computations, absolute by
Section~\ref{subsec:windows}; the verification script in the repository
recomputes all invariants of this pair, including the per-witness Euler
characteristic identity, in seconds. Minimality is the exhaustive
$n \le 4$ statement above.
\end{proof}

\begin{remark}
$\overline{C_5}$ carries the magnitude homology of the $5$-cycle in the
sense of Hepworth--Willerton (a symmetric digraph's quasimetric is the
graph metric). The twin is not a circulant: a circulant digraph
$\mathrm{Circ}(5; S)$ with $|S| = 2$ has either zero or five digons
(according as $S \cap (-S)$ is empty or all of $S$), never three.
\end{remark}

\begin{remark}\label{rem:digons}
All three witness classes of Theorem~\ref{thm:beyonddist} pair digraphs
with different digon counts ($5$ versus $3$ in each), so the
reciprocity-aware $\dwlp$ separates each of them, and exhaustively on
$n \le 5$, no $(\dwlp, \text{distance})$-equivalent pair is separated by
$\MH$ ($k \le 3$) at all. One size up this changes:
\end{remark}

\begin{theorem}\label{thm:beyonddwlp}
There exist digraphs on six vertices, minimally, that are
$\dwlp$-equivalent and have equal multisets of directed distances but
are separated by $\MH$ and by $\PH$. Witness (machine-verified by two
independent implementations, exact): the digraphs
\texttt{\&EEDBaHS} and \texttt{\&EEDB`KQ} of
Appendix~\ref{app:witnesses}. Both have twelve arcs, three digons,
out-degree $2$ at every vertex, in-degree sequence $(1,1,1,3,3,3)$,
identical distance multisets ($12 \times 1$, $12 \times 2$,
$6 \times 3$), and both are strongly connected; but
$\MH_{2,3} = \QQ^6$ versus $0$, and
$\PH_\ast = (\QQ, \QQ, 0)$ versus $(\QQ, 0, \QQ^2)$. Exactly $24$ such
$(\dwlp, \text{distance})$-equivalent classes on six vertices are split
by $\MH$; none exist on at most five vertices.
\end{theorem}

\begin{proof}
Finite checks and exact machine computations, as in
Theorem~\ref{thm:beyonddist}; every reported witness was re-verified
with the independent dense implementation and structural (digest-free)
$\dwlp$ comparison. Minimality is the exhaustive $n \le 5$ statement of
Remark~\ref{rem:digons}.
\end{proof}

\begin{remark}\label{rem:roundtripdepth}
The mechanism (Section~\ref{sec:mech}) identifies what is being
detected, and it is finer than counting. The first witness has return
distances $(2,2,2,3,3,3)$ on its six one-way arcs, the three return-$2$
arcs forming the directed triangle $3 \to 4 \to 5 \to 3$; the second
has return distance $2$ on all six. Both therefore abound in
asymmetric length-$3$ round trips, but in the second witness every
such round-trip configuration is pinned in the sense of
Definition~\ref{def:midpointgraph} (geodesically fillable), so its
diagonal contribution to $\MH_{2,3}$ vanishes, while the one-way
triangle of the first survives unpinned (three diagonal plus three
off-diagonal classes). Two things follow. First, what degree-two
magnitude homology sees beyond $(\dwlp, \text{distances})$ is
round-trip structure \emph{modulo geodesic fillability}, not any raw
count of reciprocated arcs. Second, the return-distance profiles of
the two witnesses differ, so the arc-typing refinement $\dwl^{r}$ of
Question~\ref{q:returntyped} does separate this pair; whether it
subsumes the degree-two signal everywhere is exactly that question.
\end{remark}

\section{The mechanism: degree-two magnitude homology}\label{sec:mech}

The separations of Theorems~\ref{thm:incomparable}, \ref{thm:dwlplus},
\ref{thm:beyonddist}, and \ref{thm:beyonddwlp} all occur in magnitude
homology of degree two.
This section explains why, using a component-counting description of
$\MH_{2,l}$. A complete basis description of the second magnitude
homology of a digraph was recently given by Ivanov and Mukoseev
\cite{IvanovMukoseev24} (long paths modulo short congruence); the
midpoint-graph formulation below is an equivalent form suited to our
purposes, with a short self-contained proof, and the corollaries drawn
from it, which concern separating power, do not appear there.

The magnitude chain complex splits over ordered endpoint pairs: the
differential never deletes endpoints, so
$\MC_{k,l}(G) = \bigoplus_{(a,b)} \MC_{k,l}(a,b)$, where
$\MC_{k,l}(a,b)$ is spanned by the tuples starting at $a$ and ending at
$b$ (with $a = b$ allowed).

\begin{definition}\label{def:midpointgraph}
Fix a digraph $G$, an ordered pair $(a,b)$, and $l \ge 2$. The
\emph{midpoint graph} $\Gamma_l(a,b)$ has vertex set
$V_l(a,b) = \{\, y \notin \{a,b\} : d(a,y) + d(y,b) = l \,\}$ (both
distances finite). Every $4$-tuple $(a,x,y,b) \in \MC_{3,l}(a,b)$
(so $x \ne a$, $x \ne y$, $y \ne b$, total length $l$) is classified by
the two interior face conditions
\[
\mathrm{(A)}\; d(a,x) + d(x,y) = d(a,y), \qquad
\mathrm{(B)}\; d(x,y) + d(y,b) = d(x,b):
\]
if both hold it contributes an \emph{edge} $\{x, y\}$ of
$\Gamma_l(a,b)$; if exactly one holds it contributes a \emph{pin} on
the surviving midpoint ($y$ for (A), $x$ for (B)); if neither holds it
contributes nothing. (When a condition holds, the surviving midpoint
automatically lies in $V_l(a,b)$.)
\end{definition}

\begin{proposition}\label{prop:mh2structure}
For every digraph $G$ and $l \ge 2$,
\[
\dim \MH_{2,l}(G) \;=\;
\sum_{(a,b)\,:\,d(a,b) = l} \bigl(c(\Gamma_l(a,b)) - 1\bigr)
\;+\;
\sum_{(a,b)\,:\,d(a,b) \ne l} c^{\circ}(\Gamma_l(a,b)),
\]
where $c$ is the number of connected components (isolated vertices
included) and $c^{\circ}$ is the number of components containing no
pinned vertex. Moreover, when $d(a,b) = l$ no pins occur.
\end{proposition}

\begin{proof}
Work in one summand $(a,b)$. The complex is
$\MC_{3,l}(a,b) \to \MC_{2,l}(a,b) \to \MC_{1,l}(a,b)$ with
$\MC_{2,l}(a,b) = \QQ V_l(a,b)$ and $\MC_{1,l}(a,b) = \QQ$ if
$d(a,b) = l$ and $0$ otherwise. The boundary of a $4$-tuple
$(a,x,y,b)$ is $-[\mathrm{A}]\,(a,y,b) + [\mathrm{B}]\,(a,x,b)$, i.e.,
in the basis $V_l(a,b)$: an edge column $\pm(e_x - e_y)$ if both
conditions hold, a single $\pm e$ column (a pin) if one does, zero if
none does. The rank of a matrix whose columns are edge differences and
pins on a vertex set $V$ equals $|V'| - c_{\mathrm{np}}$, where $V'$ is
the set of touched vertices and $c_{\mathrm{np}}$ the number of touched
components without pins; hence its corank on $\QQ V$ is the number of
components (isolated vertices included) without pins, which is
$c^{\circ}(\Gamma_l(a,b))$.

If $d(a,b) \ne l$ the outgoing differential is zero, so
$\dim\MH_{2,l}(a,b)$ equals that corank. If $d(a,b) = l$, then first,
every $y \in V_l(a,b)$ maps to $-(a,b)$, so the outgoing differential
has rank $1$ and kernel the sum-zero hyperplane, of dimension
$|V| - 1$; second, $\partial^2 = 0$ on a $4$-tuple evaluates to
$([\mathrm{B}] - [\mathrm{A}])\,(a,b)$, forcing
$[\mathrm{A}] = [\mathrm{B}]$: no pins occur, and the image is spanned
by edge differences, of rank $|V| - c$. Hence
$\dim\MH_{2,l}(a,b) = (|V| - 1) - (|V| - c) = c - 1$.
\end{proof}

Beyond the proof above, Proposition~\ref{prop:mh2structure} was
verified mechanically against the general implementation on all
$9{,}842$ digraphs with $3 \le n \le 5$, in every grading, and
Corollary~\ref{cor:mh22} additionally on all $1{,}540{,}944$ digraphs
with $n = 6$, with zero mismatches in every case.

\begin{corollary}\label{cor:mh22}
For every digraph,
\[
\dim \MH_{2,2}(G) \;=\; \sum_{v} \deg^{-}(v)\deg^{+}(v)
\;-\; \#\{(x,z) : d(x,z) = 2\}.
\]
Consequently $\MH_{2,2}$ is determined by round-one D-WL data together
with the distance multiset; in particular the beyond-distance
separations of Theorem~\ref{thm:beyonddist} cannot occur in bidegree
$(2,2)$, and indeed all three witness classes separate first at
$(2,3)$.
\end{corollary}

\begin{proof}
At $l = 2$ there are no $4$-tuples ($\MC_{3,2} = 0$), so
$\Gamma_2(a,b)$ has no edges and no pins, and every component is an
isolated vertex. The pairs with $d(a,b) = 2$ contribute
$m(a,b) - 1$ each, where $m(a,b) = |V_2(a,b)|$ is the number of
midpoints; the remaining pairs contribute $|V_2(a,b)|$. Summing,
$\dim\MH_{2,2} = \sum_{(a,b)} |V_2(a,b)| - \#\{d(a,b) = 2\}$, and
$\sum_{(a,b)} |V_2(a,b)|$ counts all composable arc pairs
$x \to y \to z$ ($x \ne y \ne z$, $x = z$ allowed), which is
$\sum_y \deg^{-}(y)\deg^{+}(y)$. The first term is determined by the
multiset of degree pairs, which round-one D-WL refines; the second by
the distance multiset. The statement about the witnesses is read off
the computed tables (Table~\ref{tab:star}).
\end{proof}

\begin{corollary}\label{cor:cycles}
For the directed cycle $C_n$ ($n \ge 3$):
$\dim \MH_{2,n}(C_n) = n$, supported entirely on the diagonal pairs
$(a,a)$.
\end{corollary}

\begin{proof}
For $a \ne b$ and $l = n$: a midpoint $y$ has
$d(a,y) + d(y,b) \in \{d(a,b),\, d(a,b) + n\}$, never $n$, so
$V_n(a,b) = \emptyset$. For $a = b$: every $y \ne a$ satisfies
$d(a,y) + d(y,a) = n$, so $V_n(a,a)$ is all $n{-}1$ other vertices. A
$4$-tuple $(a,x,y,a)$ has total length $n$ exactly when $x$ strictly
precedes $y$ in the cyclic order from $a$, and then both face
conditions hold (each is an identity of cyclic gaps), so
$\Gamma_n(a,a)$ is a complete graph: one component, no pins, and
$d(a,a) = 0 \ne n$ gives a contribution of $c^{\circ} = 1$ per vertex
$a$.
\end{proof}

\begin{remark}\label{rem:girth}
Corollary~\ref{cor:cycles} is a directed-girth phenomenon: degree-two
magnitude homology detects the length of shortest directed cycles, in
analogy with the undirected girth results of Asao, Hiraoka, and
Kanazawa \cite{AHK24}. It is the mechanism behind
Theorem~\ref{thm:dwlplus}: $C_3 \sqcup C_3$ has directed girth $3$ and
$\MH_{2,3} = \QQ^6$ (each triangle contributes its three diagonal
classes), while $C_6$ has directed girth $6$ and $\MH_{2,3} = 0$, its
degree-two homology appearing only at $l = 6$. Refinement is blind to
this because both digraphs are locally identical ($1$-in-$1$-out
everywhere); the round trips that magnitude homology counts are exactly
the correlation it cannot see.
\end{remark}

\begin{remark}\label{rem:twinsmech}
The same lens explains the beyond-distance witnesses of
Theorem~\ref{thm:beyonddist}. For the twin pair, the machine
decomposition of $\MH_{2,3}$ by endpoint pairs (script
\texttt{src/mh2.py}) gives: for the symmetric $\overline{C_5}$, all
ten classes sit on off-diagonal pairs at distance $2$ and the diagonal
contributes $0$ (every round trip has even length); for the twin,
the diagonal contributes $2$, generated by its four one-way arcs
(arcs with $d(x,y) = 1$, $d(y,x) = 2$: asymmetric round trips of
length $3$), and off-diagonal pairs contribute $4$. Diagonal summands
of $\MH_{2,l}$ count components of odd or asymmetric round-trip
configurations, i.e., reciprocity structure: precisely the correlation
that D-WL (which reads in- and out-neighborhoods separately) and the
distance multiset both discard.
\end{remark}

\section{The landscape on at most six vertices}\label{sec:landscape}

\begin{table}[t]
\centering
\caption{Number of equivalence classes induced on the isomorphism classes
of digraphs by each invariant (machine-generated). ``dist'' is the
multiset of ordered directed distances; ``all three'' is the joint
$(\MH_{k\le3}, \PH_{\le 2}, \RH_{\le 3})$ class.}
\label{tab:classes}
\tabclasscounts
\end{table}

\begin{table}[t]
\centering
\caption{Refinement failures: number of equivalence classes of the first
invariant that split into two or more classes of the second
(machine-generated). A zero row is an exhaustive determination statement
for that $n$ and those windows.}
\label{tab:splits}
\tabsplits
\end{table}

Tables~\ref{tab:classes} and \ref{tab:splits} summarize the exhaustive
sweep. Beyond the theorems above, two features are worth recording.

First, D-WL is nearly complete on this range ($9{,}567$ classes out of
$9{,}608$ at $n = 5$; $1{,}539{,}145$ out of $1{,}540{,}944$ at
$n = 6$, and $\dwlp$ reaches $1{,}540{,}606$), consistent with the
classical almost-sure completeness of color refinement; the homology
theories are far coarser as graph-level summaries ($75{,}221$ $\MH$
classes, $21$ $\PH$ classes, $14$ $\RH$ classes at $n = 6$). Their
value is therefore not raw separating power but the incomparability
established above: what they see, refinement can miss entirely.

Second, within the MPSS family: $\PH$ does not determine $\RH$ (one
class splits at $n = 4$, three at $n = 5$, nine at $n = 6$), but $\MH$
and $\PH$ \emph{jointly} determine $\RH$ throughout the range, in the
computed windows; internally, the joint $(\MH, \PH, \RH)$ partition at
$n = 5$ has exactly two more classes than the $\MH$ partition, and both
extra splits are the $E^1$-versus-$E^2$ witnesses of
Section~\ref{sec:e1e2}. The $E^1$-versus-$E^2$ phenomenon itself
becomes abundant one size up: $1{,}631$ $\MH$-classes split by $\PH$ at
$n = 6$.

\begin{question}\label{q:mhphrh}
Do the bigraded ranks of $\MH$ together with $\PH_\ast$ determine
$\RH_\ast$ for all digraphs? (True for all $1{,}550{,}786$ digraphs on
at most six vertices in the computed windows.)
\end{question}

\begin{question}\label{q:returntyped}
Let $\dwl^{r}$ denote directed refinement with every arc typed by the
directed distance from its head back to its tail (its return distance;
$\dwlp$ is the coarsening that only distinguishes return distance $1$
from $> 1$). Does $(\dwl^{r}, \text{distances})$ subsume degree-two
magnitude homology, or does the pattern of
Theorems~\ref{thm:beyonddist} and \ref{thm:beyonddwlp} continue at some
larger $n$?
\end{question}

\subsection*{Methodology and verification}

All digraphs were enumerated with nauty (\texttt{geng | directg})
\cite{nauty}; counts match OEIS A000273 ($16$, $218$, $9{,}608$,
$1{,}540{,}944$ for $n = 3, \dots, 6$). All boundary matrices are
integer matrices and all ranks are computed exactly over $\QQ$; no
floating point is used anywhere. Two independent implementations exist
(a dense Fraction-based one and a sparse fraction-free one sharing no
linear algebra code) and agree on all $218$ digraphs with $n = 4$ and a
$400$-digraph sample at $n = 5$; the degree-two formulas of
Section~\ref{sec:mech} were additionally verified against the general
implementation on all $9{,}842$ digraphs with $n \le 5$ in every
grading, and the $\MH_{2,2}$ formula on all $1{,}540{,}944$ digraphs
with $n = 6$. The $n = 6$ sweep ($25$ minutes on nine worker processes)
stores canonically digested refinement histories; a digest collision
could only merge two classes, never separate equal ones, so no reported
separation can be an artifact, and every witness pair claimed in this
paper was re-verified from scratch with the independent dense
implementation and digest-free structural refinement comparison ($42$
pairs, all confirmed). The
implementation is validated by a test suite of $69$ checks against
independently known values: the graded Euler characteristic of $\MH$
against the magnitude power series (computed independently from the
similarity matrix) across all gradings $l \le 6$ on seven digraphs and
graphs; Hepworth--Willerton diagonality and ranks for complete graphs;
$\partial^2 = 0$ for the regularized path boundary on all $218$
digraphs with $n = 4$; the standard path homology examples of
\cite{GLMY12} (directed cycles, filled and unfilled triangles and
squares); and contractibility statements for reachability homology. The
star witnesses of each theorem were additionally re-verified in isolation
with extended windows ($k \le 5$, $l \le 10$ for magnitude homology) and
per-witness Euler characteristic checks.

\section{Discussion}\label{sec:discussion}

\paragraph{Implications for directed graph learning.}
The invariants of the MPSS are neither stronger nor weaker than the
refinement bound governing standard directed message passing: they
separate D-WL-equivalent digraphs already at four vertices (and even
$\dwlp$-equivalent ones at six), while missing distinctions (weak
connectivity, degree profiles) that one round of refinement sees. For
feature design the conclusion is symmetric: homological features are
candidates to add genuinely new signal to directed MPNNs, exactly as
persistent homology is for undirected GNNs \cite{BallesterRieck25}, but
they cannot replace refinement. The mechanism section sharpens
this: what magnitude homology adds over refinement-plus-distances is, in
degree two, exactly the directed round-trip structure counted by the
diagonal summands of Proposition~\ref{prop:mh2structure} (girth and
reciprocity), which standard D-WL provably misses and which practical
architectures only partially recover by edge-typing reciprocal arcs:
Theorem~\ref{thm:beyonddwlp} shows that unpinned round-trip structure
survives even that typing, and Question~\ref{q:returntyped} asks for the
refinement level at which the degree-two signal is finally absorbed.

\paragraph{Implications within the MPSS.}
Theorem~\ref{thm:e1e2} shows in the strongest rank-level sense that the
$E^1$ differential of the MPSS carries information: two digraphs can have
identical $E^1$ rank functions and different $E^2$ pages, indeed
different limits. The layered family of Lemma~\ref{lem:sourcesink} may be
of independent interest as the locus where $E^1$ degenerates to counting
while $E^2$ recovers the topology of the underlying graph.

\paragraph{Limitations.}
Exhaustive claims are limited to $n \le 6$ and to the stated windows
(equalities only; inequalities are absolute). We compare against D-WL and
$\dwlp$, not against the higher levels of a directed WL hierarchy; a
directed analogue of the $k$-WL placement of \cite{BallesterRieck25}, and
the extension from unfiltered digraph homology to persistent path
homology \cite{ChowdhuryMemoli18} and persistent magnitude homology
\cite{BiFengLiWu26} on weighted digraphs, are natural next steps. Our
lower bounds transfer: any invariant that specializes to $\MH$, $\PH$, or
$\RH$ at a fixed filtration value inherits the corresponding
separations.

\paragraph{Data and code availability.}
The full implementation (exact-arithmetic magnitude, path, and
reachability homology of digraphs; D-WL and $\dwlp$; enumeration and
analysis drivers), all sweep outputs, and scripts reproducing every
number and table in this paper are available at
\url{https://github.com/krishnatheaverage/mpss-dwl-expressivity} and
archived at Zenodo (DOI \texttt{10.5281/zenodo.21138733}; all versions
\texttt{10.5281/zenodo.21138732}).

\section*{Declarations}

\noindent\textbf{Funding.}\ No funding was received for conducting this
study.

\begin{appendices}

\section{Witness data}\label{app:witnesses}

Digraphs are given in nauty digraph6 format together with arc lists on
vertex set $\{0, \dots, 4\}$; $uv$ denotes the arc $u \to v$. The three
D-WL-and-distance-equivalent classes on five vertices split by magnitude
homology (Theorem~\ref{thm:beyonddist}) are represented by:

\beyonddistpairs

\medskip
In each pair the first digraph has five digons and the second has three.
The first pair is the star witness of Theorem~\ref{thm:beyonddist}; in
the remaining two pairs the path homology dimensions $(\PH_0, \PH_1,
\PH_2)$ agree, so the separation there is by magnitude homology alone.

\medskip
The witness pair of Theorem~\ref{thm:beyonddwlp}
($\dwlp$-equivalent, distance-equivalent, on vertex set
$\{0, \dots, 5\}$):

\medskip
\texttt{\&EEDBaHS}: arcs $\{03, 04, 13, 15, 24, 25, 30, 34, 42, 45,
51, 53\}$; digons $\{0,3\}$, $\{1,5\}$, $\{2,4\}$; one-way arcs $04$,
$13$, $25$ with return distance $3$ and $34$, $45$, $53$ (a directed
triangle) with return distance $2$.

\texttt{\&EEDB`KQ}: arcs $\{03, 04, 13, 15, 24, 25, 30, 35, 42, 43,
51, 54\}$; digons $\{0,3\}$, $\{1,5\}$, $\{2,4\}$; all six one-way
arcs $04$, $13$, $25$, $35$, $43$, $54$ have return distance $2$.

\end{appendices}

\bibliography{refs}

\end{document}
